A neural coarse graining theory of consciousness/\newline Non-trivial Informational closure through neural coarse-graining as the Fundamental Informational Correlates of Consciousness

### 2019-04-25 sections/4/section-4-main.tex
Language *before* 4.1; commented out Writing Materials

% SECTION 
% 	A neural coarse graining theory of consciousness

\rlstart{5}

        In this section, we propose a new theoretical framework of consciousness \{\tnamefull\}\todo{We need a good name}. This framework proposes that NTIC is the fundamental property for information processing to be conscious. Additionally, coarse-graining is necessary to achieve NTIC in various of different systems.

        We starts our argument from the first presumptive fact that every human conscious precept is highly informative. Our conscious experience always consists of high multidimensional information and usually involves information from multiple modalities. The notion of information referred to in this article is considered as resolved uncertainty. Being a certain conscious state rules out other enormous possible conscious experience in the state space of conscious experience. Therefore, every conscious percept resolves a huge amount of uncertainty and provides considerable information.

        This notion is in agreement with one of the axiom, \textit{information}, in Integrated Information Theory (IIT 3.0) which claims that \myQuote{...an experience of pure darkness is what it is by differing, in its particular way, from an immense number of other possible experiences...} \citep[p. 2]{oizumi2014phenomenology}.

        Based on this presumptive fact, we postulate that the amount of information possessed of a conscious percept should be equivalent to information involved in a particular event or entity in the physical reality. Strictly speaking, we recognise information is a common language between conscious experience and the physical reality. We further take a radical standpoint and postulate that consciousness is fundamentally associated with information, i.e. information correlates of consciousness (ICC), \cite{chalmers1996conscious, tononi2004information, gamez2011information, Gamez2016}) and discard the notion of neural correlates of consciousness(NCC).

        Next, we next consider the second presumptive fact that the information in consciousness is bound. Even though information involved in every conscious percept is extremely rich, it is still limited and we do not consciously feel everything in this universe. This leads to the boundary problems of consciousness \cite{goff2006experiences}. There should exist information boundaries for every conscious entity. Presuming consciousness is information-based phenomenon, we can identify the boundaries with measures from information theory. Particularly, in this case, boundaries can be defined by information closure. Therefore, information closure should be observed in conscious entities in the physical reality.

        \rewording{informational closure also suggests that there is no information flow from the lower to the higher level. Knowledge of the microscopic levels will not improve predictions of the future states of the macroscopic levels \citep{PFANTE.2014}}. 
        Information closure also implies that the system is ignorant of information outside of the system. Therefore, if a coarse-grained level achieves information closure, it suggests the NTIC process at the coarse-grained level is ignorant of information processing happening at microscopic levels and also higher coarse-grained levels. This virtually makes the informational closed macroscopic levels as a complete reality. 
        \todo{Due to my concern mentioned last week, I think I should also mention the closure is not only to the environment but also to the microscopic levels}

@@

        \begin{WritingMaterials}
        For example, considering two physical systems (Alice and Bob) and there is no interaction (or information passing) in between. Assuming that Alice is conscious, the conscious experience that Alice has should not involve any information about Bob.

        \todo{Considering putting this para}
        Similarly, we can apply this scenario to brain processing between two regions. We can consider a thought experiment. 
        Imagining that in the future, one poor subject attend this experiment. We remove the subject's visual cortex from other brain areas. However, we keep all the connection between the two parts of the brain intact by replacing dendrites and axons between the two parts with micro electronic cables. Because these cables are very long, we can move the visual cortex to the surface of Mars. Now the experimenters give visual cortex a input signal. If the signal travels with the speed of light, it still takes three seconds to reach the other part on the earth. We can ask: Between the experimenters send the input signal and the information reach the part on the earth, what visual conscious content does the subject experience in the three seconds? We presume that in this three seconds subject's conscious experience should not contain any information about the visual input. Otherwise, the subject may be able to use the information to perform subsequent actions. This violates physics. This may look like a trivial questions.

        However, in many systems, it is difficult to achieve NTIC at microscopic levels due to stochasticity. Through coarse-graining, NTIC processes become possible at macroscopic levels. Therefore, coarse-graining is necessary for many biological systems to implement NTIC processes, especially for the neural system (Fig. \ref{fig:LevelOfConsciousness}). In the following, we explained this framework and provide mathematical definitions and implications of ICC.  
        \end{WritingMaterials}

		\begin{WritingMaterials}

        @ As we mentioned in Chapter 3, to form non-trivial informational closure has a huge evolutionary advantage for the creatures interacting with the environment at our scale.

        @ The physical environment dynamic at the human scale are nearly deterministic.\needref{Do we need ref here?}

		@ To maximise fitness for individual human-being, neural system should try hard to infer and model the deterministic physical rules.

		@ This drives the neural system to create a non-trivial informational closure.

        @ Therefore, NTIC emerges from evolutionary processes is functionally plausible.

        @ Importantly, at the microscopic level, it is impossible to achieve NTIC as we mentioned in chapter 2.

		@ However, at a macroscopic scale, information processing can achieve non-trivial informational closure (Fig. \ref{fig:LevelOfConsciousness})

        @ Therefore, the NTIC correlating conscious experience must settled at coarse-grained macroscopic levels.

		@ Through coarse-graining, we can define a new set of variables.

        @ information is scale-dependent

		@ If the new set of variables are informational closure, then this creates a new \textit{reality} for all the variables inside.        
		@ \textit{Reality} here means that the all the future states of the system are driven by the current information (include variables and the relationships) within the boundary.

        @

        @ Relationship between coarse-graining, scale and interaction with environment

		@ \rewrite{
			This illustrates how coarse-graining allows the formulation of incomplete generalizations that are relatively invariant, although at the cost of predictive precision regarding fine-grained details.} \cite{price2007causation}

			@@ conscious perception are very stable and time invariant.
			@@ Conscious perception is informative but we loss all the precise information about the states at microscopic level.

        @ To summaries, in our framework, consciousness is 
			    @@ Intrinsic

			    @@ Informative

				@@ scale-dependent

        % our claim   
			@ We claim that the state of a coarse-grained scale which realises non-trivial informational closured correlates the state of consciousness.

			@ We further postulate that level of consciousness corresponds to informational closureness of processes.

			@ The contents of consciousness correspond to the states of the non-trivial informational closure.

		% Following
			@ In the following part, we....\todo{todo}

			@ This is the same as "if a tree falls in a forest" thought experiment.

			@ Therefore, to understand this system, you don't need more information from external environment.

			@ It also means everything in this system is self-defined.

			@ \critical{I need to somehow acknowledge \cite{pennartz2017consciousness}}
				@@ \rewrite{ In conclusion, for a well-behaved representational system, it does not matter whether a neural activation pattern represents an external world feature truthfully or not, as long as the system represents the feature consistently over time and in sensory space.}

				@@ pennartz levels are defined by functional elements call "functional ensembles". \cite{pennartz2017consciousness} \rewrite{This delineation is different from a scale-based distinction of levels, because different levels are distinguished here based on function , i.e., what each level can accomplish in terms of computation and representation, culminating in perception at the highest level.}

			@ \rewrite{
				Multilevel views on consciousness and cognition have been expressed before by many other theorists, such as Oppenheim and Putnam (1958), Attneave (1961), Hofstadter (1979, 1985), Churchland and Sejnowski (1992), Wimsatt (1994), and Lycan (1996).} \cite{pennartz2017consciousness}

				@@ Oppenheim , P. , \& Putnam , H. ( 1958 ). Unity of science as a working hypothesis . In H. Feigl , G. Maxwell , \& M. Scriven (Eds.), Concepts, theories, and the mind – body problem (pp. 3 – 36 ). Minneapolis : University of Minnesota Press .
				@@ Attneave , F. ( 1961 ). In defense of homunculi . In W. A. Rosenblith (Ed.), Sensory communication (pp. 777 – 782 ). Cambridge, MA : MIT Press .
				@@ Hofstadter , D. R. ( 1979 ). Godel, Escher, Bach, an eternal golden braid . New York : Basic Books .
				@@ Churchland , P. S. , \& Sejnowski , T. J. ( 1992 ). The computational brain . Cambridge, MA : MIT Press .
				@@ Wimsatt , W. C. ( 1994 ). The ontology of complex systems: levels of organization, perspectives, and causal thickets. Canadian Journal of Philosophy , 20 , 207 – 274 .
				@@ Lycan ,  W. G.  ( 1996 ).  Consciousness and experience .  Cambridge, MA :  MIT Press
			@ It's important that not the neural states but the coarse-grained state determine contents of consciousness.

@@

\rlstart{3}
		\subsection{Measure of consciousness/information correlate of consciousness}	
		Based on our core assumption that NTIC is the core informational entity correlating conscious experience, we propose conscious levels and contents can be derived by computes the properties of a NTIC process. \newline

		\noindent
		We first hypothesise that, at given time $t$, levels of consciousness $C_{t}^{Level}$ corresponds to the degree of  $NTIC_{t}$ at a coarse-grained level.
			\begin{equation}\label{eq:cLevel}
				C_{t}^{Level} = NTIC_{t}
			\end{equation}
		\newline
		\noindent 
		We second hypothesise that, the content of consciousness corresponds to the state of the NTIC process.
			\begin{equation}\label{eq:cContent}
				C_{t}^{Content} = Y_{t}
			\end{equation}

		We discuss the implications of the two hypotheses in the following. 	
	    \subsection{Level of Consciousness correlates degrees of NTIC of a process}

>>>>

% SECTION 
% 	A neural coarse graining theory of consciousness
\rlstart{5}
		In this section, we propose a new theoretical framework of consciousness called \{\tnamefull\}  % The Neural Corase-Graining Theory of Consciousness
			\todo{We need a good name}. 
		The Theory makes two propositions. Firstly, a NTIC process sufficiently provides content of consciousness \sout{, endowing such content with an information-theortic account}. Secondly, a necessary condition to achieve NTIC is for information processing to satisfy coarse-graining.

		%
		In what follows, we will first establish two presumptions on which our Theory depends. Then, we will develop our propositions.
			\todo{
				In what specific ways are these presumptions help develop the Theory?
			}

		%
		% Presumtion I: Informativeness of conscious percepts
		The first presumption is that (human) conscious percepts are highly informative. Such informativeness holds true if conscious percepts serve to resolve sensory uncertainty. Resolved uncertainty, therefore, represents the sort of information that stands out amongst those giving rise to alternative forms of perceptual experience. This notion is in agreement with one of the axioms of the Integrated Information Theory (IIT 3.0), i.e., the \textit{information} axiom, which claims that \myQuote{\ldots an experience of pure darkness is what it is by differing, in its particular way, from an immense number of other possible experiences\ldots} \citep[p. 2]{oizumi2014phenomenology}.

		%
		Based on this presumption, we postulate that the amount of information possessed of a conscious percept should be equivalent to information involved in a particular event or entity in the physical reality.
			\todo[inline]{
				This statement seems somewhat unclear. Suggestion: information processed of a conscious percept has a physical underpinning(?)
			}\noindent
		Strictly speaking, we recognise information as a common language between conscious experience and the physical reality. We further take the standpoint and postulate that consciousness is fundamentally associated with information. Namely, we will entertain information correlates of consciousness (ICC), \cite{chalmers1996conscious, tononi2004information, gamez2011information, Gamez2016}) in place of a more restrictive notion of neural correlates of consciousness (NCC).

		%
		% Presumption II: Boundedness of conscious percepts
		Next, we consider a second presumption that the amount of
			\todo[inline]{
				What the term `boundary' refers to could be made clearer. 
				It is quite ambiguous as to whether it means limits in quantity or physical extent where 
				information is available to our sensorium as well as consciousness.
			}\noindent
		information held in consciousness is bounded. Even though information involved in every conscious percept is extremely rich, it is nevertheless limited as we are only consciously aware of a subset of events physics can plausibly realise. This speaks to the boundary problems of consciousness \cite{goff2006experiences}. The problem states that there should exist information boundaries for every conscious entity. Assuming consciousness is an information-theoretic phenomenon, we can identify a formal account \sout{and metric} on such boundary. As will be outlined later in our Theory, a boundary can be characterised by information closure. By prediction, physical entities that follow information closure are conscious.
			\todo[inline]{
				The original sentence is hard to follow: Therefore, information closure should be observed in conscious entities in the physical reality.
			}\noindent

		%
		\rewording{
			informational closure also suggests that there is no information flow from the lower to the higher level. 
			Knowledge of the microscopic levels will not improve predictions of the future states of the macroscopic levels \citep{PFANTE.2014}
		}. 
		Information closure also implies that the system is ignorant of information outside of the system. 
		Therefore, if a coarse-grained level achieves information closure, it suggests the NTIC process at the coarse-grained level is ignorant of information processing happening at microscopic levels and also higher coarse-grained levels. 
		This virtually makes the informational closed macroscopic levels as a complete reality. 
		\todo{Due to my concern mentioned last week, I think I should also mention the closure is not only to the environment but also to the microscopic levels}

@@

		\begin{WritingMaterials}
		For example, consider two physical systems (Alice and Bob) that have no interaction (or information passing) in between. Assuming that Alice is conscious, Alice's conscious experience should not involve any information about Bob.

		\todo{Considering putting this para}
		Similarly, we can conduct a thought experiment by drawing an analogy between this example and information-passing involving two brain regions. Suppose we have a participant who is to have their visual cortex surgically separated from the rest of the brain, yet keeping all the connectivity perfectly intact by means of artifical wiring. Without altering the normative functioning of the separated visual cortex, it is then sent to the surface of Mars under sufficiently long wires. We know for a fact that signals travel at the speed of light on these special wires. Still, it takes three seconds to travel end-to-end. We ask: what would our participant experience visually during the interval between the experimenter administering stimulation and that stimulation reaching earth? We conjecture that the participant's conscious experience should not contain any information about the visual stimulation. Otherwise, the participant may be able to use the information to perform subsequent actions, thus violating laws of physics. This is a seemingly trivial conclusion. 
			\todo[inline]{
				Transitioning into NTIC seems abrupt.
			}\noindent
		However, in many systems, it is difficult to achieve NTIC at microscopic levels due to stochasticity. Through coarse-graining, NTIC processes become possible at macroscopic levels. Therefore, coarse-graining is necessary for many biological systems to implement NTIC processes, especially for the neural system (Fig. \ref{fig:LevelOfConsciousness}). In the following, we explained this framework and provide mathematical definitions and implications of ICC.  
			\todo[inline]{
				Thought experiment does not appear to have an obvious connection with NTIC and coarse-graining?
			}\noindent
		\end{WritingMaterials}

		% \begin{WritingMaterials}
		% @ As we mentioned in Chapter 3, to form non-trivial informational closure has a huge evolutionary advantage for the creatures interacting with the environment at our scale. 
		% @ The physical environment dynamic at the human scale are nearly deterministic.\needref{Do we need ref here?}
		% @ To maximise fitness for individual human-being, neural system should try hard to infer and model the deterministic physical rules.
		% @ This drives the neural system to create a non-trivial informational closure.
		% @ Therefore, NTIC emerges from evolutionary processes is functionally plausible. 
		% @ Importantly, at the microscopic level, it is impossible to achieve NTIC as we mentioned in chapter 2.
		% @ However, at a macroscopic scale, information processing can achieve non-trivial informational closure (Fig. \ref{fig:LevelOfConsciousness})
		% @ Therefore, the NTIC correlating conscious experience must settled at coarse-grained macroscopic levels.
		% @ Through coarse-graining, we can define a new set of variables.
		% @ information is scale-dependent
		% @ If the new set of variables are informational closure, then this creates a new \textit{reality} for all the variables inside.
		% @ \textit{Reality} here means that the all the future states of the system are driven by the current information (include variables and the relationships) within the boundary. 
		% @ Relationship between coarse-graining, scale and interaction with environment
		% @ \rewrite{
		% 	This illustrates how coarse-graining allows the formulation of incomplete generalizations that are relatively invariant, although at the cost of predictive precision regarding fine-grained details.} \cite{price2007causation}
		% 	@@ conscious perception are very stable and time invariant.
		% 	@@ Conscious perception is informative but we loss all the precise information about the states at microscopic level.        
		% @ To summaries, in our framework, consciousness is 
		% 		@@ Intrinsic 
		% 		@@ Informative  
		% 		@@ scale-dependent 
		% % our claim   
		% 	@ We claim that the state of a coarse-grained scale which realises non-trivial informational closured correlates the state of consciousness.
		% 	@ We further postulate that level of consciousness corresponds to informational closureness of processes.
		% 	@ The contents of consciousness correspond to the states of the non-trivial informational closure.
		% % Following
		% 	@ In the following part, we....\todo{todo}
		% 	@ This is the same as "if a tree falls in a forest" thought experiment.
		% 	@ Therefore, to understand this system, you don't need more information from external environment.
		% 	@ It also means everything in this system is self-defined.
		% 	@ \critical{I need to somehow acknowledge \cite{pennartz2017consciousness}}
		% 		@@ \rewrite{ In conclusion, for a well-behaved representational system, it does not matter whether a neural activation pattern represents an external world feature truthfully or not, as long as the system represents the feature consistently over time and in sensory space.}
		% 		@@ pennartz levels are defined by functional elements call "functional ensembles". \cite{pennartz2017consciousness} \rewrite{This delineation is different from a scale-based distinction of levels, because different levels are distinguished here based on function , i.e., what each level can accomplish in terms of computation and representation, culminating in perception at the highest level.}
		% 	@ \rewrite{
		% 		Multilevel views on consciousness and cognition have been expressed before by many other theorists, such as Oppenheim and Putnam (1958), Attneave (1961), Hofstadter (1979, 1985), Churchland and Sejnowski (1992), Wimsatt (1994), and Lycan (1996).} \cite{pennartz2017consciousness}
		% 		@@ Oppenheim , P. , \& Putnam , H. ( 1958 ). Unity of science as a working hypothesis . In H. Feigl , G. Maxwell , \& M. Scriven (Eds.), Concepts, theories, and the mind – body problem (pp. 3 – 36 ). Minneapolis : University of Minnesota Press .
		% 		@@ Attneave , F. ( 1961 ). In defense of homunculi . In W. A. Rosenblith (Ed.), Sensory communication (pp. 777 – 782 ). Cambridge, MA : MIT Press .
		% 		@@ Hofstadter , D. R. ( 1979 ). Godel, Escher, Bach, an eternal golden braid . New York : Basic Books .
		% 		@@ Churchland , P. S. , \& Sejnowski , T. J. ( 1992 ). The computational brain . Cambridge, MA : MIT Press .
		% 		@@ Wimsatt , W. C. ( 1994 ). The ontology of complex systems: levels of organization, perspectives, and causal thickets. Canadian Journal of Philosophy , 20 , 207 – 274 .
		% 		@@ Lycan ,  W. G.  ( 1996 ).  Consciousness and experience .  Cambridge, MA :  MIT Press
		% 	@ It's important that not the neural states but the coarse-grained state determine contents of consciousness.
		% \end{WritingMaterials}

@@

% 
% 
% 
\rlstart{3}
		\subsection{Measure of consciousness/information correlate of consciousness}	
		Based on our core assumption that NTIC provides content of conscious experience, we express formally that the content and level of consciousness can be derived by computing the properties of a NTIC process. 
			\todo[inline]{
				What does level mean here? Level as in degree or intensity?
			}

		We first hypothesise that, at given time $t$, levels of consciousness $C_{t}^{Level}$ corresponds to the degree of $NTIC_{t}$ at a coarse-grained level.
			\begin{equation}\label{eq:cLevel}
				C_{t}^{Level} = NTIC_{t}
			\end{equation}
		%
		We second hypothesise that, the content of consciousness corresponds to the state of the NTIC process.
			\begin{equation}\label{eq:cContent}
				C_{t}^{Content} = Y_{t}
			\end{equation}

		We discuss the implications of the two hypotheses in the following.

		\subsection{Level of Consciousness correlates degrees of NTIC of a process}
